Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-12 main.tex
sec.VI

\newcommand{\graphof}[1]{\llbracket#1\rrbracket}
\newcommand{\compl}[1]{#1^\complement}
%\newcommand{\sublist}[1]{\set{#1}}
%\newcommand{\subst}[2]{#1\mathord{\scriptstyle\mapsto} #2}
\newcommand{\sublist}[1]{[#1]}
\newcommand{\subst}[2]{#1/#2}
\newcommand{\ssubst}[2]{\sublist{\subst{#1}{#2}}}

@@

A formula is
\bfit{rectified} if all bound variables are distinct from one another and from
all free variables. As every formula has a logically equivalent
rectified form, we will restrict most of the results in this paper to rectified formulas.

@@

\begin{definition}
A \bfit{first-order combinatorial proof} of a fograph $G$ is a skew
bifibration $f: N \rightarrow \gG$ where $N$ is a fonet. A \bfit{first-order
combinatorial proof} of a formula $A$ is a combinatorial proof of its graph
$\graphof{A}$.
\end{definition}

@@

@@

\subsection{The Linear Fragments $\FOMLL$ and $\FOMLS$}

>>>>

\newcommand{\rectif}[1]{\widehat{#1}}
\newcommand{\graphof}[1]{\llbracket#1\rrbracket}
\newcommand{\compl}[1]{#1^\complement}
%\newcommand{\sublist}[1]{\set{#1}}
%\newcommand{\subst}[2]{#1\mathord{\scriptstyle\mapsto} #2}
\newcommand{\sublist}[1]{[#1]}
\newcommand{\subst}[2]{#1/#2}
\newcommand{\ssubst}[2]{\sublist{\subst{#1}{#2}}}

@@

A formula is \bfit{rectified} if all bound variables are distinct from
one another and from all free variables. Every formula can be
transformed into a logically equivalent rectified form, by simply
renaming the bound variables. If we consider formulas equivalent
modulo $\alpha$-conversion (renaming of bound variables), then the
rectified form of a formula $A$ is uniquely defined, and we denote it
by $\rectif A$.

@@

\begin{definition}
A \bfit{first-order combinatorial proof} (\bfit{FOCP}) of a fograph $G$ is a skew
bifibration $f: N \rightarrow \gG$ where $N$ is a fonet. A \bfit{first-order
combinatorial proof} of a formula $A$ is a combinatorial proof of its graph
$\graphof{A}$.
\end{definition}

\begin{thm}[\cite{hughes:fopws}]
  \label{thm:FOCP}
  FOCPs are sound and complete for first-order logic.
\end{thm}

@@

Let us now establish the connection between derivations in $\FOKS$ and
combinatorial proofs.

\begin{thm}\label{thm:CP-DI}
  Let $A$ be a rectified formula, and assume we have a combinatorial proof $\phi\colon\gC\to\graphof A$. Then there is a derivation
  \begin{equation}
    \label{eq:decom}
    \vlderivation{
      \vlde{\set{\wlD,\wrD,\acD,\me,\mfaD,\mexD,\fequ}}{\Deri_2}{A}{
        \vlde{\FOMLS}{\Deri_1}{A'}{
          \vlhy{\ttt}}}}
  \end{equation}
  for some $A'$ such that $\graphof{\rectif{A'}}=\gC$. Conversely,
  whenever we have a derivation as in~\eqref{eq:decom} above, then
  there is a combinatiorial proof $\phi\colon\gC\to\graphof A$ such that $\gC=\graphof{\rectif{A'}}$.
\end{thm}

Furthermore, in the proof of Theorem~\ref{thm:CP-DI}, we will see that
(i) the links in the fonet $\gC$ correspond precicely to the pairs of
atoms that meet in the instances of the $\aiD$-rule in the derivation
$\Deri_1$, and (ii) the ''flow-graph'' of $\Deri_2$ that traces the
quatifier- and atom-occurrences in the derivation corresponds exactly
to the vertex-mapping induced by $\phi$.

Thus, combinatorial proofs are closely related to derivations of the
form~\eqref{eq:decom}, and since by Theorem~\ref{thm:decomposition}
every derivation can be transformed into that form, we can say that
combinatorial proofs form a canononical proof representation for first
order logic, similarly to what proof nets are for linear
logic~\cite{girard:96:PN}.

Finally, Theorems~\ref{thm:KS1}, \ref{thm:decomposition}
and~\ref{thm:CP-DI} imply Theorem~\ref{thm:FOCP}, which means that we
have here and alterantive proof of the soundness and competeness of
first-order combinatorial proofs which is simpler than the one given
in~\cite{hughes:fopws}.

@@

In this section we prove Theorems
\subsection{The Linear Fragments $\FOMLL$ and $\FOMLS$}
